Nathan Greenslit Data Analysis Project

### 2023-04-20 code/analysis_code/dust/dust_and_copies.qmd
Addedto env data viz
edits

```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = t7sum),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 10))

@@

```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

>>>>

```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = t7sum),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density",
    title = "Dust AOD over Daily Time Series")+
  theme(
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 20))

@@

```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
  # axis.text.x = element_blank(),
   # axis.title.x = element_blank(),
   # axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

### 2023-04-20 products/manuscript/Manuscript.qmd
edits

*What I am looking for* I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels. If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response. Nutrients, dust input, salinity, and temperature will be the primary factors of interest.

*How I will analyze it* - Look at Vibrio growth over time series (estimated from qPCR) - Examine relationships between dust input and Vibrio growth - Examine influence of site-specific environmental parameters (Temperature, Salinity, Nutrients) on growth response - Run linear models on factors like: dust x growth and nutrients x growth - NMDS to see which parameters have the strongest influence on growth

@@

A high resolution (daily) time series was conducted starting on July 7th and ending on July 19th, 2022 capturing before, during, and after a Saharan dust event. Samples were collected in 1L autoclaved Polypropylene Bottles and immediately placed in a cooler on ice. Following transport to the laboratory, the samples were concentrated onto 0.2µm pore size, hydrophobic polycarbonate membranes (25mm in diameter) and stored at -80°C until DNA Extraction using a ZymoBIOMICS DNA Miniprep Kit (Cat#:4300).

@@

Total *Vibrio* concentrations were quantified using a SYBR Green quantitative PCR (qPCR) method. Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) PCR product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3. Master mix and PCR conditions were derived from [@westrichSaharanDustNutrients2016a]. Briefly, 5µL of 2X SYBR PowerUp Master Mix (Applied Biosystems, Foster City, CA) were added to primers with a final concentration of 0.16µM and PCR water for a final volume of 10µL. Reactions were run in triplicate on a CFX96 Touch Real-Time PCR Detection System (Bio-Rad Laboratories, Hercules, CA) with the following cycling conditions: 2 minutes at 50◦C for UDG activation and 95◦C for 2 min to activate AmpliTaq polymerase and UP, followed by 40 cycles of 95◦C for 3 seconds for denaturation and 60◦C for 30 seconds for annealing and extension. Each run was followed by a dissociation step (60C to 95C by 0.5C increments) to determine a melt curve for analysis of specificity. Each test was run with positive and negative controls in triplicate. Cycle threshold values for each qPCR test was compared to a standard curve representing 10\^1 to 10\^6 gene copies per reaction volume. <!--# Add Std Curve specs -->

@@

## Exploratory/Descriptive analysis

#### Dust AOD and Vibrio Enumeration <!--# What are AOD units? -->

This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series. The daily time series captured two dust events. One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57. Dust AOD returned to lower concentrations following the time series. The top portion of the figure depicts total enumerated Vibrio as copies per mL, with color by site. There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.

![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png){width="1491"}

@@

![Figure 2:](images/000019.png)

## Statistical analysis

### Check Distributions:

@@

### Cross-correlation Analysis:

@@

### Correlation Matrices

@@

![](images/000175.png)

@@

![Figure 4](images/000055.png){width="1491"}

@@

@@

@@

While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 1). Field studies have suggested that salinities for optimal growth lie between 5-25, depending on the species [@sullivanEffectsTemperatureSalinity2018]. The higher salinities at this site (40-42) may have attributed to this observed weaker growth response.

@@

## Strengths and Limitations

At Blind Oso for example, the salinity is observed to drop from 42-20 within an eight day period that coincided with dust input. It is difficult to parse these two variables out as either or both could have contributed to growth at this site.

<!--# Difficult to parse out sites  -->

<!--# Sample size  -->

>>>>

I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels. If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response. Nutrients, dust input, salinity, and temperature will be the primary factors of interest.

@@

A high resolution (daily) time series was conducted starting on July 7th and ending on July 19th, 2022 capturing before, during, and after a Saharan dust event. Samples were collected in 1L autoclaved Polypropylene Bottles and immediately placed in a cooler on ice. Following transport to the laboratory, the samples were concentrated onto 0.2µm pore size, hydrophobic polycarbonate membranes (25mm in diameter) and stored at -80°C until DNA Extraction using a ZymoBIOMICS DNA Miniprep Kit (Cat#:4300). To evalulate contamination in the lab, negative controls (DI and MilliQ Water) were introduced at sample collection, filtration, DNA Extraction, and downstream quantitative polymerase chain reaction (qPCR).

@@

Total *Vibrio* concentrations were quantified using a SYBR Green qPCR method. Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) PCR product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3. Master mix and PCR conditions were derived from [@westrichSaharanDustNutrients2016a]. Briefly, 5µL of 2X SYBR PowerUp Master Mix (Applied Biosystems, Foster City, CA) were added to primers with a final concentration of 0.16µM and PCR water for a final volume of 10µL. Reactions were run in triplicate on a CFX96 Touch Real-Time PCR Detection System (Bio-Rad Laboratories, Hercules, CA) with the following cycling conditions: 2 minutes at 50◦C for UDG activation and 95◦C for 2 min to activate AmpliTaq polymerase and UP, followed by 40 cycles of 95◦C for 3 seconds for denaturation and 60◦C for 30 seconds for annealing and extension. Each run was followed by a dissociation step (60C to 95C by 0.5C increments) to determine a melt curve for analysis of specificity. Each test was run with positive and negative controls in triplicate. Cycle threshold values for each qPCR test was compared to a standard curve representing 10\^1 to 10\^6 gene copies per reaction volume.

@@

## Exploratory Analysis

### Visualization

#### Dust AOD <!--# What are AOD units? -->

This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series. The daily time series captured two dust events. One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57. Dust AOD returned to lower concentrations following the time series.

![Figure 1:](images/000009-01.png)

This figure depicts total enumerated Vibrio as copies per mL, with color by site. There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.

![Figure 2:](images/000005-03.png)

@@

![Figure 3:](images/000019.png)

### Statistical analysis

#### Check Distributions:

@@

#### Cross-correlation Analysis:

@@

#### Correlation Matrices

@@

![Figure 4:](images/000175.png)

@@

![Figure 5:](images/000055.png){width="1491"}

@@

![Figure 6:](images/000017-02.png)

![Figure 7:](images/000007-02.png)

@@

##

@@

While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 2). Field studies have suggested that salinities for optimal growth lie between 5-25, depending on the species [@sullivanEffectsTemperatureSalinity2018]. The higher salinities at this site (40-42) may have attributed to this observed weaker growth response.

@@

## Limitations

One primary limitation of this project was the small samples size (n = 12 per site). One sample per day limits the interpretation of the results of this project and changes the scope from an explanatory report to exploratory.

Additionally, while qPCR can reliably detect the concentration of DNA of the target, the results do not differentiate between living and dead cells, and therefore provides an estimate of quantification.

While it can be proposed that the degree of microbial response is contingent on the environmental settings of the site, further work is needed to look at each specific environmental parameter and how it may impact *Vibrio* abundance. At Blind Oso for example, the drastic drop in salinity during an eight day period coincided with dust input. It then becomes difficult to parse these two variables out as either may have contributed to growth at this site.
